---
title: "Heterogeneous cost and budgets"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Heterogeneous cost and budgets}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.width = 7, fig.height = 5)
```

```{r setup}
library(paretoinfer)
```

Evaluating an alternative is rarely free, and rarely equally expensive across
alternatives: a heuristic with more restarts, a model with more parameters, a
field trial at a remote site. The design records a cost per alternative, the
next-evaluation rule divides uncertainty by cost, and a budget stops the
identification with an honest "inconclusive" when it runs out.

```{r problem}
means <- rbind(A1 = c(0.15, 0.70), A2 = c(0.40, 0.40), A3 = c(0.70, 0.15),
               A4 = c(0.42, 0.43), A5 = c(0.60, 0.60), A6 = c(0.30, 0.80))
simulate <- function(alternative, n, h = 0.1) {
  mu <- means[alternative, ]
  data.frame(y1 = runif(n, mu[1] - h, mu[1] + h), y2 = runif(n, mu[2] - h, mu[2] + h))
}
```

## A budget that runs out

`A5` and `A6` cost five units per evaluation, the others one. With a budget
of 400 units the procedure stops before anything is certified.

```{r budget}
design <- pareto_design(alternatives = rownames(means), objectives = c("y1", "y2"),
                        bounds = c(0, 1), alpha = 0.05, epsilon = 0.05,
                        cost = c(1, 1, 1, 1, 5, 5), budget = 400)
set.seed(2)
state <- pareto_run(design, simulate, batch = 5, initial = 5)
state
plausible_pareto_set(state)$table[, c("alternative", "status", "n_evaluations", "cost")]
```

The expensive alternatives received only their initial evaluations: the rule
preferred to narrow the cheap boxes. The outcome is inconclusive, with the
estimated frontier as the point estimate and every alternative still
plausible. The report says so and records the budget.

```{r report}
pareto_report(state)
```

## Continuing with a larger budget

A stopped identification accepts no further evaluations, so that a report
always describes a design that was fixed in advance. To continue, a new
design with a larger budget is run; the guarantee is per identification, and
the sampling of the first run cannot be reused without a new design that
includes it.

```{r larger}
design2 <- pareto_design(alternatives = rownames(means), objectives = c("y1", "y2"),
                         bounds = c(0, 1), alpha = 0.05, epsilon = 0.05,
                         cost = c(1, 1, 1, 1, 5, 5), budget = 4000)
set.seed(2)
state2 <- pareto_run(design2, simulate, batch = 5, initial = 5)
glance(state2)[, c("n_evaluations", "total_cost", "n_dominated", "identified", "stopping_reason")]
plausible_pareto_set(state2)$table[, c("alternative", "status", "dominated_by", "n_evaluations", "cost")]
```

`A5` and `A6` were certified dominated with few evaluations: a clearly
dominated alternative needs little evidence, whatever it costs.

## Driving the loop by hand

When evaluations come from outside R, or when the stopping rule is not the
design's budget but a judgement call, the loop is written by the user.
`choose_next_evaluation()` proposes, `update_objectives()` consumes, and
`plausible_pareto_set()` can be inspected at any time without invalidating
anything.

```{r manual}
design3 <- pareto_design(alternatives = rownames(means), objectives = c("y1", "y2"),
                         bounds = c(0, 1), alpha = 0.05, epsilon = 0.05, cost = c(1, 1, 1, 1, 5, 5))
set.seed(3)
state3 <- initialize_pareto(design3)
for (k in design3$alternatives) {
  ev <- simulate(k, 5); ev$alternative <- k
  state3 <- update_objectives(state3, ev)
}
repeat {
  sets <- plausible_pareto_set(state3)
  if (sets$identified || state3$total_cost >= 1500) break
  nxt <- choose_next_evaluation(state3)
  ev <- simulate(nxt$alternative, 5); ev$alternative <- nxt$alternative
  state3 <- update_objectives(state3, ev)
}
state3
```

Stopping by hand at any time keeps every certificate valid: the guarantee is
simultaneous over time, so no choice of when to look or when to stop can
break it. What the user gives up by stopping early is only the size of the
plausible set.
